---
title: Analyzing Himalayan climbing expeditions
author: Alex Cookson
date: '2020-04-01'
slug: analyzing-himalayan-climbing-expeditions
categories: []
tags: []
description: ''
topics: []
---  
\  

In this series of posts, we will analyze climbing expeditions to the Himalayas, a mountain range comprising over 50 mountains, including [Mount Everest](https://en.wikipedia.org/wiki/Mount_Everest), the highest mountain (above sea level) in the world.  
\  

This is *Part 2* of a three-part series:

- [Part 1 looked at Himalayan peaks and their first ascents](https://www.alexcookson.com/post/analyzing-himalayan-peaks-first-ascents/)
- Part 2 (this post) looks at Everest expeditions
- Part 3 will look at Himalayan expeditions overall (excluding Mount Everest) *(coming soon)*  
\  

This post will focus on expeditions to Mount Everest, the tallest mountain in the world and the most popular destination for Himalayan expeditions.
\  


## Setup

First, we'll load the `tidyverse` and some other useful packages:

- `fishualize` for ggplot2 colour palettes [inspired by fish](https://cran.r-project.org/web/packages/fishualize/vignettes/overview_colors.html)
- `scales` for turning ugly numbers into pretty numbers (e.g., 0.3 to 30%)
- `extrafont` for additional fonts to use in graphs

We'll also set a default `ggplot2` theme and import our data.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(fishualize)
library(scales)
library(extrafont)

theme_set(theme_light())

peaks <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/peaks.csv")
expeditions <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/expeditions.csv")
members <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/members.csv")
```  
\  


## Data inspection

We have two data tables to investigate: `expeditions`, which has data on entire climbing expeditions, and `members`, which has data on the individual members of those expeditions.

Let's start with **expeditions**.

```{r}
expeditions %>%
  glimpse()
```

`expeditions` has four different types of data:

1. Basic data on the peak. We could bring in the `peaks` table if we wanted information like alternative names or heights.
2. Important expedition dates -- when they reached basecamp, when they reached the summit, and when the expedition ended
3. "Outcomes", including elevation highpoint and the reason for the expedition ending
4. Miscellaneous additional data, like number of members, deaths, and whether oxygen was used  
\  

What about expedition **members**?

```{r}
members %>%
  glimpse()
```

`members` has NUMBER types of data:

1. Some of the same data from the `peaks` and `expeditions` tables, here for convenience
2. Demographic data, like age and sex
3. Individual outcomes, like elevation high and macabre "outcomes" like injury and death
4. Miscellaneous additional data, like expedition role and whether the individual used oxygen





Number of expeditions

```{r}
expeditions %>%
  count(peak_id, peak_name, name = "expeditions") %>%
  mutate(peak_name = fct_reorder(peak_name, expeditions)) %>%
  top_n(20, wt = expeditions) %>%
  ggplot(aes(expeditions, peak_name, fill = peak_name)) +
  geom_col() +
  scale_fill_fish(option = "Pseudocheilinus_tetrataenia",
                  discrete = TRUE) +
  theme(legend.position = "none",
        text = element_text(family = "Bahnschrift"))
```

Create Everest datasets

```{r}
everest_expeditions <- expeditions %>%
  filter(peak_name == "Everest")

everest_members <- members %>%
  filter(peak_name == "Everest")
```
















## Conclusion

CONCLUSION TEXT

If you enjoyed this post, [let me know](https://twitter.com/alexcookson) on Twitter and read the other posts analyzing this dataset:

- [Part 1 looked at Himalayan peaks and their first ascents](https://www.alexcookson.com/post/analyzing-himalayan-peaks-first-ascents/)
- Part 3 will look at Himalayan expeditions overall (excluding Mount Everest) *(coming soon)*







## Appendix

Expeditions:

- What peaks were the most popular to climb?
- How big are expeditions?
- How many people are typically hired?
- What are the success rates of expeditions?
- Why do expeditions fail?
- What is the threat rate for injuries and deaths?
- Are there any particularly dangerous peaks?
- What proportion of expeditions use oxygen?


Members:

- What's the typical composition of expedition members? Age, sex, citizenship?
- Are any particular mountains more popular with certain countries?
- Are there any particularly dangerous expedition roles
- What types of injuries and causes of death are there?
- Are there any particularly dangerous elevations?
- Are younger climbers more likely to attempt solo climbs or not use oxygen?
