---
title: Analyzing Himalayan climbing expeditions
author: Alex Cookson
date: '2020-04-10'
slug: analyzing-himalayan-climbing-expeditions
categories: []
tags: []
description: ''
topics: []
---  
\  

In this series of posts, we will analyze climbing expeditions to the Himalayas, a mountain range comprising over 50 mountains, including [Mount Everest](https://en.wikipedia.org/wiki/Mount_Everest), the tallest mountain in the world.  
\  

This is *Part 3* of a three-part series:

- [Part 1 looked at Himalayan peaks and their first ascents](https://www.alexcookson.com/post/analyzing-himalayan-peaks-first-ascents/)
- [Part 2 looked at how dangerous it is to climb Mount Everest](https://www.alexcookson.com/post/how-dangerous-is-climbing-mount-everest/)
- Part 3 (this post) looks at Himalayan expeditions overall (excluding Mount Everest) 
\  

This post will focus on expeditions that *didn't* go to Mount Everest. Specifically we'll answer these questions:

- Q1?
- Q2?
- Q3?


## Setup

First, we'll load the `tidyverse` and some other useful packages:

- `fishualize` for ggplot2 colour palettes [inspired by fish](https://cran.r-project.org/web/packages/fishualize/vignettes/overview_colors.html)
- `scales` for turning ugly numbers into pretty numbers (e.g., 0.3 to 30%)
- `extrafont` for additional fonts to use in graphs
- `ggtext`, which "provides rich-text (basic HTML and Markdown) support for ggplot2" (from the package's [GitHub](https://github.com/wilkelab/ggtext))

We'll also set a default `ggplot2` theme and import our data.

```{r setup, warning = FALSE, message = FALSE}
library(tidyverse)
library(fishualize)
library(scales)
library(extrafont)
library(ggtext)

theme_set(theme_light())

peaks <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/peaks.csv")
expeditions <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/expeditions.csv")
members <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/members.csv")
```  
\  

We already have a good sense of what the data contains from earlier posts in this series, so we won't inspect the data. If you're reading this out of order, I encourage you to start with [Part 1](https://www.alexcookson.com/post/analyzing-himalayan-peaks-first-ascents/) or [Part 2](https://www.alexcookson.com/post/how-dangerous-is-climbing-mount-everest/).  
\  


## What do people climb -- other than Everest?

TEXT

```{r expeditions-by-peak, eval = FALSE}
expeditions %>%
  inner_join(peaks %>% select(peak_id, height_metres), by = "peak_id") %>%
  filter(peak_name != "Everest") %>%
  mutate(eight_thousander = height_metres >= 8000) %>%
  count(peak_id, peak_name, eight_thousander, name = "expeditions") %>%
  mutate(peak_name = fct_reorder(peak_name, expeditions)) %>%
  top_n(20, wt = expeditions) %>%
  ggplot(aes(expeditions, peak_name, fill = eight_thousander)) +
  geom_col() +
  scale_x_continuous(labels = label_comma()) +
  scale_fill_manual(values = c("#6baed6", "#08519c")) +
  labs(title = "<span style = 'color:#08519c'>**8000ers**</span> and <span style = 'color:#6baed6'>**other**</span> Himalayan mountains are both popular",
       subtitle = "Number of expeditions from 1905-2019",
       caption = "Source: The Himalayan Database",
       x = "",
       y = "") +
  theme(
    legend.position = "none",
    text = element_text(family = "Bahnschrift"),
    plot.title = element_markdown(),
    axis.text = element_text(size = 10)
  )
```



## Conclusion


If you're curious about these (or other questions!), please use this data to look into them yourself and [let me know](https://twitter.com/alexcookson) on Twitter.

And, if you enjoyed this post, I encourage you to read the other posts analyzing this dataset:

- [Part 1 looked at Himalayan peaks and their first ascents](https://www.alexcookson.com/post/analyzing-himalayan-peaks-first-ascents/)
- [Part 2 looked at how dangerous it is to climb Mount Everest](https://www.alexcookson.com/post/how-dangerous-is-climbing-mount-everest/)







## Appendix

Expeditions:

- How big are expeditions?
- How many people are typically hired?
- What are the success rates of expeditions for different peaks?
- Why do expeditions fail?
- What is the threat rate for injuries and deaths?
- Are there any particularly dangerous peaks?
- What proportion of expeditions use oxygen?


Members:

- What's the typical composition of expedition members? Age, sex, citizenship?
- Are any particular mountains more popular with certain countries?
- Are there any particularly dangerous expedition roles?
- What types of injuries and causes of death are there?
- Are there any particularly dangerous elevations?
- Are younger climbers more likely to attempt solo climbs or not use oxygen?
