---
title: Analyzing Himalayan climbing expeditions
author: Alex Cookson
date: '2020-03-22'
slug: analyzing-himalayan-climbing-expeditions
categories: []
tags: []
description: ''
topics: []
---  
\  

Introduction


## Setup

First, we'll load the `tidyverse` and import our data.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)

theme_set(theme_light())

peaks <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/peaks.csv")
expeditions <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/expeditions.csv")
members <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/himalayan-expeditions/members.csv")
```  
\  


## Data inspection

Let's take a look.

```{r}
peaks %>%
  ggplot(aes(height_metres)) +
  geom_histogram(binwidth = 100)
```


## Appendix

### Preliminary Questions

Peaks:

- How tall are Himalayan peaks?
- When were they first climbed?
- What countries were involved in first ascents?


Expeditions:

- What peaks were the most popular to climb?
- How big are expeditions?
- How many people are typically hired?
- What are the success rates of expeditions?
- Why do expeditions fail?
- What is the threat rate for injuries and deaths?
- Are there any particularly dangerous peaks?
- What proportion of expeditions use oxygen?


Members:

- What's the typical composition of expedition members? Age, sex, citizenship?
- Are any particular mountains more popular with certain countries?
- Are there any particularly dangerous expedition roles
- What types of injuries and causes of death are there?
- Are there any particularly dangerous elevations?
- Are younger climbers more likely to attempt solo climbs or not use oxygen?
