---
title: "What do countries' diplomatic gifts tell us about them?"
description: "Text analysis of diplomatic gifts to U.S. federal employees"
author: "Alex Cookson"
date: "2020-05-19"
output: html_document
---  
\  

Have you ever brought a bottle of wine, flowers, or [chocolate babka](https://www.youtube.com/watch?v=i78azsi7M94) to a dinner party as a host/hostess gift? Or brought home a souvenir for your parents, partner, or kids after you've been travelling -- like chocolate from Switzerland or, uh... [Brazil nuts](https://www.youtube.com/watch?v=tKXEydXbuOU) from Brazil? Countries do the same thing, kind of.

[Diplomatic gifts](https://diplomacy.state.gov/diplomacy/why-do-diplomats-give-gifts/) are often exchanged when dignitaries travel abroad or receie visitors. They can be lavish, like a \$780,000 emerald and diamond jewellery set, given by King Abdullah of Saudi Arabia. They can be modest, like a \$50 bottle of wine, given by President Stjepan Mesić of Croatia. Or they can be essentially worthless, but symbolic, like a live shamrock, given by Prime Minister Bertie Ahern of Ireland. (Though in the shamrock's case, the State Department *did* value it at \$5.)

**Do countries tend to give certain types of gifts?** A shamrock is meaningful coming from an Irish Prime Minister, but it would feel less appropriate if it came from, say, the Prime Minister of Japan.

We're going to perform a **text analysis** of diplomatic gifts to U.S. federal employees, which includes folks like the President, Vice President, and Secretary of State. By law, these gifts must be described, valued, and recorded. All this information is published publicly, which is what we'll use as data to answer our question.  
\  


## Setup

First, we'll load the `tidyverse` and some other useful packages:

- `tidytext` for text analysis
- `fishualize` for beautiful ggplot2 colour palettes
- `extrafont` for additional fonts
- `ggtext` for rich-text customization of text (e.g., changing colours of individual words in a plot title)  
\  

Then we'll import our data, originally from the [U.S. Office of the Federal Register](https://www.archives.gov/federal-register), which provides access to federal government documents, including the reports of diplomatic gifts that we're looking for. The original documents needed extensive cleaning to get into a [tidy format](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html). I won't go into detail on that process here, but if you're interested, look at the [README](https://github.com/tacookson/data/tree/master/us-government-gifts) on GitHub.

```{r setup-and-import, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
library(fishualize)
library(extrafont)
library(ggtext)

theme_set(theme_light())

gifts <- read_csv("https://raw.githubusercontent.com/tacookson/data/master/us-government-gifts/gifts.csv")
```  
\  


## Data inspection

Let's unwrap `gifts`!

```{r inspect-data}
# glimpse bottom of data because top doesn't have good examples
gifts %>%
  arrange(-id) %>%
  glimpse()
```

There are four types of data:

1. Who **received** the gift, including name/title and what agency they are associated with
2. Who **gave** the gift, including name/title and the country they represent
3. Details on the **gift itself**, mostly in the written `gift_description`, but also in the `value_usd` field, which gives the estimated value
4. Details about the `occasion of gifting`, like when it was received and the reason the gift was accepted  
\  

We'll focus on the `gift_description` field, which has *wonderful* free-form text descriptions. (I spent hours reading some of these descriptions.) Here is a short example of a gift given in 2003 to First Lady Laura Bush by French President Jacques Chirac.

```{r example-gift-description}
gifts %>%
  filter(id == 1605) %>%
  pull(gift_description)
```

(We'll soon see from our analysis that *Hermès* items are a quintessentially French gift.)  
\  


## Text analysis with TF-IDF

These descriptions are great, but how do we use them to answer our question? In [*Text Mining with R*](https://www.tidytextmining.com/tfidf.html), David Robinson and Julia Silge note that "a central question in text mining and natural language processing is how to quantify **what a document is about**" (emphasis mine). One way we can do that is by looking at the individual words that make up a "document" -- in this case, a gift description. If a word (like *Hermès*) shows up a lot in descriptions from a certain country (like *France*), then we can make the argument that that word is representative of that country's gifts. It gets more complicated than that, but for now let's see if we can split up these descriptions into the individual words that comprise them -- a process known as *tokenizing*. We'll use `unnest_tokens()` from the `tidytext` package.

### Tokenizing gift descriptions

```{r gift-tokens-with-stop-words}
gift_words_raw <- gifts %>%
  # Select unique gift descriptions to pare down gifts that often take up multiple lines
  distinct(gift_description, .keep_all = TRUE) %>%
  transmute(id,
            donor_country,
            gift_description = str_to_lower(gift_description)) %>%
  # Filter out travel, which isn't interesting
  filter(!str_detect(gift_description, "travel")) %>%
  # Break gift_description into individual words
  unnest_tokens(output = word, input = gift_description)
```  
\  

What does the *tokenized* dataset look like?

```{r tokenized-description-example}
gift_words_raw %>%
  head(11)
```

Each word has its own row, but we can still read the beginning of the first description, a gift from Moldova: "40 x 29 gilt framed oil painting of an autumn landscape".

What if we stopped now and looked at the most common words used in gifts from Moldova?

```{r example-with-stop-words}
gift_words_raw %>%
  filter(donor_country == "Moldova") %>%
  count(word, sort = TRUE) %>%
  head(12)
```

Not very insightful. The most common word is "of", which doesn't tell me anything about distinctively *Moldovan* gifts. And it looks like there are some other words that probably appear in most or all descriptions. For example, "est" and "value" are probably common because description need to included an est(imated) value!  
\  


### Eliminating stop words

What are stop words and what will we be using:

- Normal stop words
- Month stop words
- Country stop words
- Description stop words

```{r stop-words}
# Country stop words (e.g., Afghani, Croatian, Tanzanian)
source("https://raw.githubusercontent.com/tacookson/data/master/us-government-gifts/src/country-adjective-list.R")

# Stop words specific to this dataset
gift_stop_words <- c("x", # Used for describing dimensions (e.g, 6" x 6")
                     "inches",
                     "est", # Common words about value and disposition
                     "rec'd",
                     "recd",
                     "disposition",
                     "administration",
                     "services",
                     "pending",
                     "transfer",
                     "transferred",
                     "archives",
                     "protocol",
                     "gsa",
                     "national",
                     "foreign",
                     "records",
                     "de", # Stop word in romance languages like French/Spanish
                     "la",
                     "los")

# Month name stop words
month_stop_words <- str_to_lower(c(month.abb, month.name))

# Put all the stop words together
custom_stop_words <- tibble(stop_word = c(country_stop_words,
                                          gift_stop_words,
                                          month_stop_words)) %>%
  unnest_tokens(word, stop_word)
```

```{r gift-words-cleaned}
gift_words <- gift_words_raw %>%
  # Get rid of standard English stop words (or, the, an, etc.)
  anti_join(stop_words, by = "word") %>%
  # Get rid of our custom stop words
  anti_join(custom_stop_words, by = "word") %>%
  # Only keep words that have at least one letter
  # Numbers only aren't likely meaningful in our context
  filter(str_detect(word, "[a-z]"),
         !is.na(donor_country))
```


### Calculating TF-IDF

Decision to require that gifts be given at least 3 times

```{r tf-idf}
gift_words_tfidf <- gift_words %>%
  distinct(id, donor_country, word) %>%
  count(word, donor_country) %>%
  bind_tf_idf(word, donor_country, n) %>%
  # Only include words that appears in at least 3 gift descriptions
  filter(n >= 3)
```

```{r}
gift_words_tfidf %>%
  filter(donor_country == "Kenya") %>%
  arrange(desc(tf_idf))
```



### Overall top words

Explanation of `ggtext`? Or just use it?

```{r top-tf-idf-overall}
# Set defaults for annotations so we don't need to copy-paste them
# fill = "white" doesn't work for some reason
update_geom_defaults(GeomRichtext, list(hjust = 0, colour = "grey20", label.colour = NA,
                                        family = "Bahnschrift", size = 4))

gift_words_tfidf %>%
  top_n(16, wt = tf_idf) %>%
  mutate(word = fct_reorder(str_to_title(word), tf_idf)) %>%
  ggplot(aes(tf_idf, word, fill = donor_country)) +
  geom_col() +
  geom_text(aes(label = donor_country),
            col = "white",
            fontface = "bold",
            hjust = 1,
            nudge_x = -0.005) +
  # Iceland
  geom_richtext(x = 0.307, y = 13.5, fill = "white",
                label = "Iceland often gifts the book set<br>\"The <span style = 'color:#669638'>**Complete Sagas**</span> of <span style = 'color:#669638'>**Icelanders**</span>\"") +
  # Marshall Islands
  geom_richtext(x = 0.248, y = 10.5, fill = "white",
                label = "Marshall Islands gifted many <span style = 'color:#BCA51F'>**elephant tusks**</span><br>in 2001, which were purposefully <span style = 'color:#BCA51F'>**destroyed**</span>") +
  # Sweden
  geom_richtext(x = 0.202, y = 9, fill = "white",
                label = "Sweden gifts <span style = 'color:#F0A630'>**Orrefors**</span> crystal") +
  # Ghana
  geom_richtext(x = 0.161, y = 7, hjust = 0, fill = "white",
                label = "Ghana gifts <span style = 'color:#92B7B7'>**Kente**</span>, a traditional Ghanian fabric") +
  # Cuba
  geom_richtext(x = 0.138, y = 1, hjust = 0, fill = "white",
                label = "Cuba, naturally, gifts <span style = 'color:#99AFD6'>**cigars**</span>") +
  scale_fill_fish("Epibulus_insidiator", discrete = TRUE) +
  scale_colour_fish("Epibulus_insidiator", discrete = TRUE) +
  expand_limits(x = c(0, 0.5)) +
  labs(title = "What are the most \"distinctive\" words associated with diplomatic gifts?",
       x = "TF-IDF",
       y = NULL) +
  theme_minimal() +
  theme(legend.position = "none",
        text = element_text(family = "Bahnschrift"),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 14),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", face = "bold"))
```

Annotation notes for above:

- Iceland likes giving the book set "The Complete Sagas of Icelanders" (id 762, 7026, 7790)
- Luxembourg likes giving fox sculptures (id 2123, 2239, 2678)
- Ghana likes giving Kente cloth gifts, which is a traditional cloth of Ghana (id 880, 2527, 8361)
- Moldova likes giving cognac (id 5932, 8293)
- Cuba likes giving cigars (id 568, 7152, 6959)
- Sweden likes giving Orrefors crystal (id 654, 1742, 6195)


### Specific countries

```{r interesting-countries}
g7_countries <- c("Italy", "Japan", "Germany", "Canada", "United Kingdom", "France")

g20_countries <- c("Argentina", "Australia", "Brazil", "Canada", "China", "France", "Germany",
                   "India", "Indonesia", "Italy", "Japan", "Mexico", "Russia", "Saudi Arabia",
                   "South Africa", "South Korea", "Turkey", "United Kingdom")
```

```{r g7-countries-tf-idf}
gift_words_tfidf %>%
  filter(donor_country %in% g7_countries) %>%
  group_by(donor_country) %>%
  top_n(6) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, by = tf_idf, within = donor_country, sep = ": ")) %>%
  ggplot(aes(tf_idf, word, fill = donor_country)) +
  geom_col() +
  scale_y_reordered(sep = ": ") +
  scale_fill_fish(option = "Epibulus_insidiator", discrete = TRUE) +
  facet_wrap(~ donor_country, scales = "free_y") +
  labs(title = "What words are associated with countries' gifts to the U.S.?") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 14),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", face = "bold"))
```

```{r g20-countries-tf-idf}
gift_words_tfidf %>%
  filter(donor_country %in% setdiff(g20_countries, g7_countries)) %>%
  group_by(donor_country) %>%
  top_n(6) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, by = tf_idf, within = donor_country, sep = ": ")) %>%
  ggplot(aes(tf_idf, word, fill = donor_country)) +
  geom_col() +
  scale_y_reordered(sep = ": ") +
  scale_fill_fish(option = "Epibulus_insidiator", discrete = TRUE) +
  facet_wrap(~ donor_country, scales = "free_y") +
  labs(title = "What words are associated with different countries' diplomatic gifts to the U.S.?") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 14),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", face = "bold"))
```


### Drawbacks of TF-IDF

Discuss shortcoming of TF-IDF approach:

- You might get words that aren't meaningful to answering your question (e.g., Moscow, Kremlin for Russia)
- Tokenizing by individual words ignores significance of compound nouns (e.g., mother-of-pearl for South Korea, leather briefcase for Argentina)
- Words used in one or two "documents" can have large TF-IDF, which can be misleading. We filtered for words that appear in at least three gift descriptions to address this, but adding that filter has its own issues because we set an arbitrary threshold and might be losing out on interesting information in words that do appear in only 1 or 2 descriptions.


## Conclusion

Conclusion goes here!



### Appendix

Questions:

- What are the most and least expensive gifts?
- Do some countries give more expensive gifts than others?
- How many of these gifts are given specifically to the President? (And what is their value?)
- What gifts has Canada given?
- Do certain countries give certain types of gifts?


Additional areas to explore:

- Looking at named entities instead of single words (would probably capture things like "men's watch" instead of separating the words)
- Stemming or lemmatization of words (e.g., Kenya had different words for "elephant" and "elephants". Those aren't really different concepts, so it would make sense to combine them into one word)
- Are there are regional patterns to gifts? (e.g., are Middle Eastern countries more likely to gift rugs than other countries?)


```{r gift-description-lookup}
gifts %>%
  filter(donor_country == "Marshall Islands",
         str_detect(str_to_lower(gift_description), "destroyed")) %>%
  View()
```