---
title: "fictional-character-personalities"
author: "Alex Cookson"
date: "11/11/2020"
output: html_document
---  
\  

Why PCA?

- PCA, dimensionality reduction
- Overview of data

I often walk through code in posts, but I won't be doing that here. Instead, I'll focus more on the analysis itself.


## Setup

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse) # For data manipulation
library(tidymodels) # For PCA
library(tidytext) # For reorder_within()
library(extrafont) # For custom fonts on graphs
library(glue) # For creating strings from multiple pieces
library(ggtext) # For fine-grained text formatting

# These are to make tables look nice on the website
# You probably won't need them
library(knitr)
library(kableExtra)

theme_set(theme_void())

personalities_raw <- read_tsv("~/data/fictional-character-personalities/personalities.txt")

genders <- read_csv("~/data/fictional-character-personalities/ref/character-gender.csv")

personalities <- personalities_raw %>%
  left_join(genders)
```  
\  

```{r data-for-pca}
pca_data <- personalities %>%
  filter(!is_emoji) %>%
  select(character_code, spectrum, mean) %>%
  pivot_wider(names_from = spectrum, values_from = mean)
```

```{r pca}
# Write recipe for PCA
pca_recipe <- recipe(~ ., data = pca_data) %>%
  # Specify character_code as key/id column
  update_role(character_code, new_role = "id") %>%
  # Normalize sets mean to zero and standard deviation of one
  step_normalize(all_predictors()) %>%
  # PCA is done here
  # Use threshold to specify we want to capture 90% of variance in the data
  step_pca(all_predictors(),
           threshold = 0.9)

# prep() implements the recipe
# bake() applies the PCA model to a dataset
personality_pca <- prep(pca_recipe) %>%
  bake(new_data = pca_data)
```

### How many principal components should I use?

- We get one component per predictor
- Our goal is dimensionality reduction, so we need to decide how many to use
- Two options: pick a number of components or pick a threshold for the amount of variance in the original data that you want the PCs to account for
- Relationship between number of components and total variance accounted for (20 components, i.e., less than 10% the number of original predictors, captures almost 90% of variance in the original data)
- In this case, I chose a threshold of 90%, which resulted in 26 PCs

```{r pca-variance-explained, fig.align = 'center', fig.width = 12, fig.height = 7.75}
# Inspect PCA to see which principal components account for what % of variance
pca_variance <- prep(pca_recipe) %>%
  tidy(type = "variance", number = 2) %>%
  filter(terms == "cumulative percent variance") %>%
  mutate(pct_variance = value / 100)

pca_variance %>%
  ggplot(aes(component, pct_variance)) +
  geom_line(size = 1.5, colour = "#2A9D8F") +
  geom_ribbon(aes(ymin = min(pct_variance), ymax = pct_variance),
              alpha = 0.3, fill = "#2A9D8F") +
  scale_x_continuous(breaks = seq(0, 240, by = 20)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1),
                     labels = label_percent(accuracy = 1)) +
  coord_cartesian(ylim = c(min(pca_variance$pct_variance), 1)) +
  labs(title = "Just a few components do a lot of heavy lifting",
       subtitle = "Cumulative variance accounted for as the number of principal components increases",
       x = "# of Principal Components",
       y = "% of Variance Accounted For") +
  theme(text = element_text(family = "IBM Plex Sans"),
        plot.title = element_text(size = 24, margin = margin(t = 10)),
        plot.subtitle = element_text(size = 15, margin = margin(t = 10, b = 10)),
        plot.margin = margin(0, 20, 20, 20),
        panel.grid.major = element_line(colour = "grey85"),
        axis.title = element_text(size = 14, colour = "grey50"),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(angle = 90, margin = margin(r = 10)),
        axis.text.x = element_text(size = 16, hjust = 0.5),
        axis.text.y = element_text(size = 16, hjust = 1))
```

```{r gender}
personalities %>%
  filter(!is_emoji,
         !is.na(gender)) %>%
  group_by(spectrum, spectrum_low, spectrum_high) %>%
  mutate(overall_avg_score = mean(mean) - 50) %>%
  ungroup() %>%
  group_by(spectrum, spectrum_low, spectrum_high, overall_avg_score, gender) %>%
  summarise(avg_score = mean(mean) - 50) %>%
  ungroup() %>%
  pivot_wider(names_from = gender, values_from = avg_score) %>%
  mutate(female = female - overall_avg_score,
         male = male - overall_avg_score) %>%
  pivot_longer(female:male, names_to = "gender", values_to = "avg_diff") %>%
  group_by(gender) %>%
  top_n(16, wt = abs(avg_diff)) %>%
  ungroup() %>%
  unite("spectrum_name", spectrum_low, spectrum_high, sep = " <---> ") %>%
  mutate(spectrum_name = reorder_within(spectrum_name, avg_diff, gender)) %>%
  ggplot(aes(avg_diff, spectrum_name, fill = gender)) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_y_reordered() +
  facet_wrap(~ gender, scales = "free_y") +
  theme_light()
```

