---
title: "sakura-senzen"
author: "Alex Cookson"
date: '`r Sys.Date()`'
output: html_document
---

Title: A step-by-step guide to `gganimate`
Subtitle: Mapping Japan's cherry blossom front

What this post does:

- Show *one possible path* from idea to beautiful animation
- Help you 

What this post does not do:

- Teach the basics of `ggplot2`
- Showcase everything `gganimate` can do


Reference `gganimate` vignette

Reference Katherine Goode's great presentation: https://goodekat.github.io/presentations/2019-isugg-gganimate-spooky/slides.html#1



### Data preparation

Goal: show the typical flowering and full-bloom dates of sakura (cherry blossom) trees in Japan. We use a dataset from (dataset provenance)

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse) # Data manipulation
library(gganimate) # Animation
library(lubridate) # Working with dates
library(ggthemes) # ggplot2 theme for maps
library(ggimage) # Images in ggplot2 plots
library(extrafont) # Custom fonts
library(ggtext) # Customising text colour in titles

library(kableExtra) # Nicely-formatted tables (you don't need this; it's just for the blog post)

sakura <- read_csv("https://raw.githubusercontent.com/tacookson/data/master/sakura-flowering/sakura-modern.csv")
```



```{r sakura-average-days}
sakura_overall <- sakura %>%
  # Ensure complete records for flowering and full-bloom dates
  filter(!is.na(flower_doy),
         !is.na(full_bloom_doy)) %>%
  # Calculate average days of year
  group_by(station_id, station_name, latitude, longitude) %>%
  summarise(flower_doy = mean(flower_doy),
            full_bloom_doy = mean(full_bloom_doy),
            .groups = "drop")

# Show what the data looks like
# (You don't need this; it's just for the blog post)
sakura_overall %>%
  head() %>%
  kable(format = "html") %>%
  kable_styling()
```

We keep day-of-year for now because numbers are easier to work with than dates. We'll fix this later.


### 1. Plot a static graph

Animations show subsets of a dataset at different times, so I like to start by showing *everything* -- all the data -- in a static graph. That way I have an overall sense of the data I'm working with. This won't make sense for some types of animations, but for this one, it's pretty helpful. The goal of our animation is to communicate the **timing** of sakura blooms in different **places**.

Let's forget the timing for now and just look at the places.

```{r static-points-only, fig.align = 'center', fig.width = 6}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude))
```

You can kind of make out the shape of Japan's islands. (Those dots at the bottom-left are smaller islands like Okinawa.)



### 2. Create the simplest animation possible

This is where we start using `gganimate` functions. To ease into things, my next step is often to create the simplest animation I can. This step I may have to pause and think -- *really think* -- about what I want to communicate to my audience, since that will dictate which `transition_*()` function to use. (`transition_*()` functions are `gganimate`'s engine. They are how we specify what data gets animated and when.)

Remember from Step 1 that we're communicating the **time** and **place** of sakura blooms. The geographic coordinates of the stations take care of the place. The animation takes care of the time.

We have two "states" of bloom. Flowering (`flower_doy`) is when 5 or 6 buds on a tree are open. And full bloom (`full_bloom_doy`) is when more than 80% of a tree's buds are open. Eventually, we'll want our animation to communication information about both states of bloom. But since we're starting simple, let's just look and when the flowering starts.

To do that, we'll use `transition_reveal()`, which starts us off without any data, adds (or, "reveals") data as time goes on, and keeps that data visible until the end of the animation. It's just like our static graph, except that individual data points will appear at different times instead all being there at once.

We need to tell `transition_reveal()` when to reveal each data point, which we do with the `along` argument. Here, we'll tell the function to reveal "along" the `flower_doy` variable. All it takes is one line of code:

```{r animation-reveal, fig.align = 'center', fig.width = 6}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude, group = station_id)) +
  transition_reveal(along = flower_doy) ###
```


### 3. Refine the transition

We've shown our audience when cherry blossoms flower, but we also want to show when they're in full bloom. To do this, we'll up the complexity (just a little!) with a different transition function: `transition_events()`. It's similar to `transition_reveal()` in that we start without data and that data gets revealed according to a variable we give.

There's one key difference, though: data won't stay visible until the end of the animation. They will **disappear** at a certain point. As with `transition_reveal()`, we just need to tell the function when the disappearance should happen. Instead of `transition_reveal()`'s `along` argument, `transition_events()` has two arguments to control the appearance and disappearance of data points: `start` and `end`.

There's one more complication. `transition_events()` also needs us to specify a window of time for the points to appear and disappear, using the `enter_length` and `exit_length` arguments. These values aren't important right now, but we need to put something in here for the animation to work. I promise we'll come back to these arguments later. For now, let's set them to 1.

```{r animation-events, fig.align = 'center', fig.width = 6}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,   ###
                    end = full_bloom_doy, ###
                    enter_length = 1,     ###
                    exit_length = 1)      ###
```



### 4. Add a dynamic title

Dynamic labels change depending on where in the animation you are. In our case, we'll add a dynamic label showing the **day of year**, since that is the scale used by our variables for bloom dates. Different `transition_*()` types have different dynamic labels. `transition_events()` uses `frame_time`.

We'll add a `labs()` argument with a `caption` argument that includes `frame_time`. We're going to wrap `frame_time` in `{curly braces}`, though, because the braces will tell the function to execute the R code within the braces. If we included `frame_time` *without* the curly braces, our title would literally show "Day of year: frame_time" and wouldn't change along with the animation.

```{r animation-points-only-add-title, fig.align = 'center', fig.width = 6}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "Day of year: {frame_time}") ###
```

Having day of year to over 10 decimal points of accuracy is distracting. However, since things inside curly braces act like R code, we can wrap the `frame_time` dynamic variable in a function and our title will show the output of that function. So we'll wrap it in a `round()` function to get it to the nearest integer.

```{r animation-points-only-add-title-rounded, fig.align = 'center', fig.width = 6}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "Day of year: {round(frame_time, 0)}") ###
```

Much better!



### 5. Add start and end pauses

The current animation is on a loop: as soon as it ends, it immediately starts the animation again from the beginning. It can be helpful to add a brief pause at the beginning and end of the animation. These pauses make it easier for our audience to tell what's going on.

To add pauses, we need to change the structure of our code. Right now, we're creating an animation that will play as soon as it finishes rendering. When we do this, `gganimate` uses some default settings, including looping the animation without any start/end pauses. We're going to switch to saving our plotting code to an object, which we'll call `p` for "plot".

Now, we can use the `animate()` function, which gives us more flexibility to tweak the settings of our animation, including start and end pauses. `animate()`'s first argument is what we want to animate, which is our plot object, `p`. This is the only required argument. If we called `animate(p)`, we would get the exact same result as we did on the previous step.

But to add pauses, we need to specify values for the `start_pause` and `end_pause` arguments. These arguments use number of frames (not seconds), which defaults to 10. Let's set both of them 15, which translates into 1.5-second pauses.

(I also add `height` and `width` arguments so that the animation is a good size for this post. These are optional if you're doing this on your own.)


```{r animation-points-only-add-pauses, fig.align = 'center', fig.width = 8}
# Assigning plot to p
p <- ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600) ###
```


### 6. Add Japan country borders

This animation is a map, so borders are a useful reference point! `ggplot2` has its own `borders()` function, which draws basic borders for common entities like countries or continents. There are more sophisticated ways of dealing with borders, but this suits our purposes just fine, so we'll stick with this simple method. We need to set the `regions` argument to "Japan".

We'll also set the `fill` and `colour` so that we have a grey theme. I'm using R's built-in greys `grey##`, where higher numbers give you a lighter grey. I know it's bland, but we'll make the colours more interesting later. Right now we're focused on getting the borders set up.


```{r animation-add-borders, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",   ###
          fill = "grey95",     ###
          colour = "grey50") + ###
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```


### 7. Use map theming

Now that we have borders, the axis labels and gridlines are getting distracting. Our audience doesn't need to pinpoint precise geographic coordinates, so let's get rid of them. The `{ggthemes}` package has a theme, `theme_map()`, that gets rid of these extras that aren't typically needed for maps.

One more thing. You probably didn't notice, but Japan is *stretched* compared to what you'd see on most maps. That's because we haven't specified a map projection to use. Let's do that with `coord_map()`, using the familiar [Mercator projection](https://en.wikipedia.org/wiki/Mercator_projection).

```{r animation-add-map-coord-and-theme, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  coord_map(projection = "mercator") + ###
  theme_map() +                        ###
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```



### 8. Add major cities as reference points

Let's add more reference points so that our audience can get their bearings. Specifically, let's add locations of major cities, like Tokyo and Kyoto. This will involve adding two elements to the animation: **points** for city locations and **text** for city names. (We'll use `geom_point` for locations and `geom_text` for names.)

We already have the data we need in `sakura_overall`:

- Geographic coordinates (`longitude` and `latitude`) tell us location
- Station names (`station_name`) are just city names, so that gives us names

But I don't want to use `sakura_overall` as the data source for our reference points. Remember, we just want to show *major* cities. If we showed every city in `sakura_overall`, it would dominate the animation and we wouldn't be able to see much else.

So let's create a separate data frame that only has the cities we want to show. I did some research to find nine cities fairly spread out across Japan, with decent-size populations. To get our data frame for these reference points, we'll add a `filter()` step that matches `station_name` to the list of cities I picked out.

```{r city-labels}
sakura_labels <- sakura_overall %>%
  filter(station_name %in% c("Sapporo", "Tokyo", "Kyoto",
                             "Hiroshima", "Niigata", "Naha",
                             "Nagasaki", "Sendai", "Hakodate")) %>%
  select(station_name, longitude, latitude)

sakura_labels %>%
  kable(format = "html") %>%
  kable_styling()
```


Now we add our `geom_point()` and `geom_text()` layers, specifying the `data` argument as `sakura_labels` (the data frame we just created).

We also added `hjust` (horizontal justification) and `vjust` (vertical justification) arguments to offset the city name from the location point. It can take some fiddling and a few tries to get the positioning exactly right. I tried 5 or 6 different combinations before I got the text positioned right where I wanted it.

```{r animation-add-cities, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,                          ###
             aes(longitude, latitude),                      ###
             size = 1) +                                    ###
  geom_text(data = sakura_labels,                           ###
            aes(longitude, latitude, label = station_name), ###
            hjust = -0.2,                                   ###
            vjust = 1) +                                    ###
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```



### 9. Replace default points with sakura icon

Using the default points to represent blooms is fine. But they're a bit boring, don't you think? Our data is about sakura trees, so let's use sakura icons! I ended up buying an icon from [Iconfinder](https://www.iconfinder.com) because it was exactly what I was looking for. But there are tons of free icons out there if you don't want to spend money.

We're going to replace our `geom_point()` function with `geom_image()`, from the `{ggimage}` package. `geom_image()` is almost the same, except that we have an extra argument -- `image` -- that we point to the image file we want to use. In my case, I save the image as `sakura-no-border.svg` in my project's main directory. If you save yours in a different place, you may need to fiddle with the file path.

Speaking of fiddling, I had to try several values for the `size` argument. You may have to do the same.

```{r animation-add-sakura-image, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,                                    ###
             aes(longitude, latitude, image = "sakura-no-border.svg"), ###
             size = 0.03) +                                            ###
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```


### 10. Convert dynamic day-of-year to calendar date

(Note: This step is a bit of a hack. If there are better ways to do it, I'd love to learn them. This method is ugly...but it works.)

Our dynamic label currently shows day-of-year, which isn't a very intuitive way to think about dates. Calendar dates are better. For example, "April 15" is more meaningful than "Day 106". So let's change our dynamic label to be a calendar date.

(WARNING: Hackiness starts here.)

My approach has two steps. First, we convert the format of `frame_time` (the dynamic label) from numeric -- which is inherited from `flower_doy` and `full_bloom_doy` -- to a date. We do this by adding the (numeric) `frame_time` to January 1. Because of how R handles addition with dates, we end up converting day-of-year to an actual calendar date.

The catch is that date objects *need* to have a year. So our January 1 "base date" needs to be January 1 of a specific. We're going to use "2019-01-01" as our base date. Why? I'm writing in 2020 and 2019 is the most recent non-leap year. (We don't want our generic calendar dates to have a February 29.) I could have just as easily used 2018, 2017, or even 1967.

I do all this conversion in curly braces in the `labs()` function's `caption` argument: pass `"2019-01-01"` into `{lubridate}`'s `ymd()` to convert text into date format, then add `frame_time`. Note that I had to add backslashes `\` before the quotation marks in `"2019-01-01"` as escape characters. If I hadn't added the backslashes, R would interpret those quotation marks as the end of the text argument of `caption` instead of *part* of the text argument. (I warned you it was hacky. Sorry :( )

How does this look?

```{r animation-convert-to-date, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "{ymd(\"2019-01-01\") + frame_time}") ###

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```

Second, we convert the [ISO 8601](https://en.wikipedia.org/wiki/ISO_8601)-formatted date to a generic calendar date, like "April 15". (Apologies to non-North American readers.) I did this by extracting the month and day from the date we calculated in step one and formatting it how I wanted:

- The `month()` function returns a number by default, so setting `label = TRUE` and `abbr = FALSE` gets us unabbreviated month names, like "April".
- The `day()` function is fine as-is. It returns a number, which is what we want.

Next, I used `paste()` to combine the two into a single string.

The date we're seeing is *technically* 2019, but since we don't show the year, it serves just fine as a generic calendar date.

```{r animation-convert-to-date-no-year, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE), day(ymd(\"2019-01-01\") + frame_time))}") ###

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```



### 11. Change font

Let's use a more interesting font. There are a couple different ways to use custom fonts in R. I used the `{extrafont}` package. Getting it set up is its own tutorial, but if you haven't used it before, the [setup instructions](https://github.com/wch/extrafont) are pretty good.

I used the [Manga Temple](https://www.dafont.com/manga-temple.font) font by [Blambot](https://twitter.com/blambot).

Once we go through `{extrafont}`'s process for using downloaded fonts, we change our `geom_text()` and `theme()` functions to include an argument that specifies Manga Temple as the font to use.

```{r animation-change-font, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            family = "Manga Temple", ###
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple")) ###

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```


### 12. Change colour palette

Let's finally get away from this drab grey colour palette. I wanted something warm that was evocative of old maps, so I went with a cream/brown palette. I got the specific hex values for colours by literally Googling "old maps" and using a colour eyedropper Chrome extension. I ended up with three colours:

- <span style = 'color:#FFF9F3'>**Cream**</span> for the background
- <span style = 'color:#AA9C8F'>**Medium Brown**</span> for Japan's landmass
- <span style = 'color:#3D1308'>**Dark Brown**</span> for points and text

We need to change colour in four places:

1. `borders()` for the landmass
1. `geom_point()` for the city location points
1. `geom_text()` for the city names text
1. `theme()` for description text (`text` argument) and background (`plot.background` argument)


```{r animation-change-colour, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F", ###
          colour = NA) +    ###
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") + ###
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308", ###
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 1,
                    exit_length = 1) +
  labs(caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"), ###
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA))    ###

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```


### 13. Tweak enter and exit animations

Now let's make the animations more nuanced. Our sakura icons are appearing a disappearing instantly, which is a bit jarring. Let's try for something more peaceful and organic, to reflect our subject. Specifically, let's mimic the blossoms themselves. Have them grow and "bloom" when they appear, then fall away when they disappear. To do this, we'll use `{gganimate}`'s `enter_*()` and `exit_*()` functions, which specify animations for data appearing and disappearing from the plot.

For the "blooming", we'll use `enter_grow()`, which starts our sakura icon small and grows to the size we specify.

For the falling away, we'll use two transitions, which will happen at the same time: `exit_shrink()`, which shrinks the icon to a small point before disappearing, and `exit_drift()`, which shifts the icon in a direction we specify before disappearing.

`exit_drift()` can take two arguments: `x_mod` for horizontal movement and `y_mod` for vertical movement. (You can combine them for diagonal movement.) Since we want our blossoms to "fall", we set `y_mod` to a negative value.

While we're at it, let's also change the timing of enter and exit animations. Previously, we gave the icons one day to enter and one day to exit. We'll do three tweaks:

1. Start the enter animation three days before the flower date
1. Give the enter and exit animations six days to do their thing. This matters now that we actually have animations that take time. Previously, we set `enter_length` and `exit_length` to 1, which didn't matter because the icons were appearing instantly.
1. Set the `range` argument. This part is new. By default, the animation starts at the lowest value in our data and stops at the highest value, but this ends us starting the animation on January 8 and ending it on May 26. Setting a custom range argument gets us "round" dates. In our case, `range = c(0, 151)` starts us on January 1 and ends us on June 1.

```{r add-growing-shrinking-transition, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy - 3, ###
                    end = full_bloom_doy,   ###
                    range(0, 151),          ###
                    enter_length = 6,       ###
                    exit_length = 6) +      ###
  enter_grow() +                            ###
  exit_drift(y_mod = -1) +                  ###
  exit_shrink() +                           ###
  labs(caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA))

animate(p, start_pause = 15, end_pause = 15, height = 715, width = 600)
```


### 14. Slow overall animation speed

Looks nice, but it feels to fast for the subject matter. I mentioned that I'm going for a peaceful, organic feel to the animation, but it feels rushed. `animate()` defaults to 100 frames at 10 FPS (frames per second), giving us a default animation of 10 seconds (plus the 1.5-second pauses we added at the start and end).

We can adjust the duration directly with the `duration` argument. We can also adjust duration indirectly with the `nframes` argument. More frames at the default 10 FPS means longer duration. I'm going to use the `nframes` method to make sure each day gets an equal number of frames. We have 152 days in the animation (remember last step when we set the range from 0 to 151?). So If we set `nframes = 152 * 2`, that will ensure each day lasts 2 frames, which is equivalent to one fifth of a second.

```{r animation-slow-speed, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA))

animate(p, nframes = 152 * 2, start_pause = 15, end_pause = 15, height = 715, width = 600) ###
```


# START HERE -- ALMOST DONE!
### 15. Add title and explanation blurb

Title text and basic formatting of size, etc.

```{r animation-add-title, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "When do cherry blossoms bloom in Japan?",                                                                    ###
       subtitle = paste("Hanami (\"flower viewing\") is a Japanese tradition over 1,000 years old, usually associated with", ###
                     "cherry blossom trees. The sakura zensen (cherry blossom front) shows the typical date of",             ###
                     "flowering and full bloom across Japan.",                                                               ###
                     "\n"),                                                                                                  ###
       sep = "\n",                                                                                                           ###
       caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA),
        plot.title = element_textbox_simple(family = "Manga Temple", colour = "#3D1308"),    ###
        plot.subtitle = element_textbox_simple(family = "Manga Temple", colour = "#3D1308"), ###
        plot.caption = element_text(family = "Manga Temple", colour = "#3D1308"))            ###

animate(p, nframes = 152 * 2, start_pause = 15, end_pause = 15, height = 715, width = 600)
```



### 16. Adjust title size and location

Very fiddly step. I often spend a lot of time on this step making sure everything is just right, through trial and error. Specifically mention `halign` and `margin` arguments.

```{r animation-adjust-title-size-location, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "/img/sakura-no-border.svg"),
             size = 0.028) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "When do cherry blossoms bloom in Japan?",
       subtitle = paste("Hanami (\"flower viewing\") is a Japanese tradition over 1,000 years old, usually associated with",
                     "cherry blossom trees. The sakura zensen (cherry blossom front) shows the typical date of",
                     "flowering and full bloom across Japan.",
                     "\n"),
       sep = "\n",
       caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA),
        plot.margin = margin(10, 50, 10, 50),                                     ###
        plot.title = element_text(halign = 0.5, margin = margin(10, 10, 10, 10)), ###
        plot.subtitle = element_text(halign = 0.5))                               ###

animate(p, nframes = 152 * 2, start_pause = 15, end_pause = 15, height = 715, width = 600)
```



### 17. Add colour to title with `ggtext`

The `<span>` tags are intimidating, but you get used to them. `ggtext` overview here: https://github.com/wilkelab/ggtext

```{r animation-add-colour-to-title, fig.align = 'center', fig.width = 8}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.028) +
  coord_map(projection = "mercator") +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  #############################################################################
  labs(title = "When do <span style = 'color:#D47FA8'>**cherry blossoms**</span> bloom in Japan?",
       subtitle = paste("<span style = 'color:#D47FA8'>**Hanami**</span> (\"flower viewing\") is a Japanese tradition over 1,000 years old, usually associated with",
                     "cherry blossom trees. The <span style = 'color:#D47FA8'>**sakura zensen**</span> (cherry blossom front) shows the typical date of",
                     "flowering and full bloom across Japan.",
                     "\n"),
  #############################################################################
       sep = "\n",
       caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA),
        plot.margin = margin(10, 50, 10, 50),
        plot.title = element_textbox_simple(size = 11, halign = 0.5, margin = margin(10, 10, 10, 10)),
        plot.subtitle = element_textbox_simple(size = 8, halign = 0.5),
        plot.caption = element_text(size = 14))

animate(p, nframes = 152 * 2, start_pause = 15, end_pause = 15, height = 715, width = 600)
```
