---
title: "sakura-senzen"
author: "Alex Cookson"
date: '`r Sys.Date()`'
output: html_document
---

Title: A step-by-step guide to `gganimate`
Subtitle: Mapping Japan's cherry blossom front

What this post does:

- Show *one possible path* from idea to beautiful animation
- Help you 

What this post does not do:

- Teach the basics of `ggplot2`
- Showcase everything `gganimate` can do


Reference `gganimate` vignette

Reference Katherine Goode's great presentation: https://goodekat.github.io/presentations/2019-isugg-gganimate-spooky/slides.html#1



### Data preparation

Goal: show the typical flowering and full-bloom dates of sakura (cherry blossom) trees in Japan. We use a dataset from (dataset provenance)

```{r setup, message = FALSE}
library(tidyverse) # Data manipulation
library(gganimate) # Plot animation
library(lubridate) # Working with dates
library(ggthemes) # ggplot2 theme for maps
library(ggimage) # Images in ggplot2 plots
library(extrafont) # Custom fonts
library(ggtext) # Customising text

library(kableExtra) # Nice tables (only needed for blog post)

sakura <- read_csv("https://raw.githubusercontent.com/tacookson/data/master/sakura-flowering/sakura-modern.csv")
```



```{r sakura-average-days}
sakura_overall <- sakura %>%
  # Ensure complete records for flowering and full-bloom dates
  filter(!is.na(flower_doy),
         !is.na(full_bloom_doy)) %>%
  # Calculate average days of year
  group_by(station_id, station_name, latitude, longitude) %>%
  summarise(flower_doy = mean(flower_doy),
            full_bloom_doy = mean(full_bloom_doy),
            .groups = "drop")

sakura_overall %>%
  head() %>%
  kable(format = "html") %>%
  kable_styling()
```

We keep day-of-year for now because numbers are easier to work with than dates. We'll fix this later.


### 1. Plot a static graph

Animations show subsets of a dataset at different times, so I like to start by showing *everything* -- all the data -- in a static graph. That way I have an overall sense of the data I'm working with. This won't make sense for some types of animations, but for this one, it's pretty helpful. The goal of our animation is to communicate the **timing** of sakura blooms in different **places**.

Let's forget the timing for now and just look at the places.

```{r static-points-only}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude))
```

You can kind of make out the shape of Japan's islands. (Those dots at the bottom-left are smaller islands like Okinawa.)



### 2. Create the simplest animation possible

This is where we start using `gganimate` functions. To ease into things, my next step is often to create the simplest animation I can. This step I may have to pause and think -- *really think* -- about what I want to communicate to my audience, since that will dictate which `transition_*()` function to use. (`transition_*()` functions are `gganimate`'s engine. They are how we specify what data gets animated and when.)

Remember from Step 1 that we're communicating the **time** and **place** of sakura blooms. The geographic coordinates of the stations take care of the place. The animation takes care of the time.

We have two "states" of bloom. Flowering (`flower_doy`) is when 5 or 6 buds on a tree are open. And full bloom (`full_bloom_doy`) is when more than 80% of a tree's buds are open. Eventually, we'll want our animation to communication information about both states of bloom. But since we're starting simple, let's just look and when the flowering starts.

To do that, we'll use `transition_reveal()`, which starts us off without any data, adds (or, "reveals") data as time goes on, and keeps that data visible until the end of the animation. It's just like our static graph, except that individual data points will appear at different times instead all being there at once.

We need to tell `transition_reveal()` when to reveal each data point, which we do with the `along` argument. Here, we'll tell the function to reveal "along" the `flower_doy` variable. All it takes is one line of code:

```{r animation-points-only}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude, group = station_id)) +
  transition_reveal(along = flower_doy)
```


### 3. Refine the `transition`

We've shown our audience when cherry blossoms flower, but we also want to show when they are in full bloom. We're going to use a more complicated function for this: `transition_events()`. This function is similar to `transition_reveal()`. We still start without data, and that data still gets revealed according to a variable we give.

There are two key differences, though. First, data won't stay visible until the end of the animation. They will **disappear** at a certain point. As with `transition_reveal()`, we just need to tell the function when the disappearance should happen.

Second, **DOES THIS FUNCTION NEED ENTER AND EXIT LENGTHS? CHECK BEFORE FINISHING THIS STEP**



we'll be using `transition_events()`, which lets us define **events**. In `gganimate`'s world, each event has its own **start** and **stop** time. This means different events can last for different periods of time. We need this for our case because different stations have different lengths of time that sakura trees are in bloom.

`transition_events()`

To do this we use the `transition_events()` function, which lets individual points stay on-screen for different periods of time. We need to specify 4 arguments:

- `start` and `end` specify when in the timeline we're creating the points appear and disappear
- `enter_length` and `exit_length` specify how long the points have to appear and disappear. These values aren't too important right now because we haven't specified the *type* of transition, so the points are appearing and disappearing instantly. But they'll become important when we start polishing our animation. The numbers use the same scale as the data, so the code below gives each point 5 days to appear and disappear. Why 5 days? No reason -- we just need to choose something and five is a nice enough number.

```{r animation-points-only}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5)
```



### 4. Add a dynamic label

Dynamic labels change depending on where in the animation you are. In our case, we'll add a dynamic label showing the **day of year**, since that is the scale used by our variables for bloom dates. Different `transition_*()` types have different dynamic labels. `transition_events()` uses `frame_time`.

We'll add a `labs()` argument with a `title` argument that includes `frame_time`. We're going to wrap `frame_time` in `{curly braces}`, though, because the braces will tell the function to execute the R code within the braces. If we included `frame_time` *without* the curly braces, our title would literally show "Day of year: frame_time" and wouldn't change along with the animation.

```{r animation-points-only-add-title}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "Day of year: {frame_time}")
```

Having day of year to over 10 decimal points of accuracy is distracting. However, since things inside curly braces act like R code, we can wrap the `frame_time` dynamic variable in a function and our title will show the output of that function. So we'll wrap it in a `round()` function to get it to the nearest integer.

```{r animation-points-only-add-title-rounded}
ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "Day of year: {round(frame_time, 0)}")
```

Much better!



### 5. Add start and end pauses

We are now assigning the plot to an object `p`, then using `p` as the first argument in the `animate()` function, which lets us tweak things like start and end pauses. Here, we'll add a pause of 20 frames to the start and end, which translates to 2 seconds with the `animate()` function's default of 10 frames per second (FPS)

```{r animation-points-only-add-pauses}
p <- ggplot() +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15)
```


Add borders of Japan.

This is an animated map, so borders will be useful reference point! `ggplot2` has its own `borders()` function, which draws basic borders for common entities like countries or continents. In our case, we'll use Japan. We'll also set the `fill` and `colour` so that we have a grey theme. A bit bland, but we'll change the colours later.

```{r animation-add-borders}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15)
```


Use map theming.

`coord_map()` translates geographic coordinates into a map projection
`theme_map()` is a clean theme that's good for maps -- eliminates a lot of the excess plot elements (e.g., gridlines, axis text)

```{r animation-add-map-coord-and-theme}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15)
```


Add cities as additional reference points.

- Create an vector of cities we want to label (city name == station_name), along with longitude and latitude
- Add tiny points for city markers
- Add city labels next to those points

Here we specify different `data` arguments in our `geom_*()` functions: vector of city labels we just created and the original data frame of sakura timing.

```{r city-labels}
sakura_labels <- sakura_overall %>%
  filter(station_name %in% c("Sapporo", "Tokyo", "Kyoto",
                             "Hiroshima", "Niigata", "Naha",
                             "Nagasaki", "Sendai", "Hakodate")) %>%
  select(station_name, longitude, latitude)

sakura_labels %>%
  kable(format = "html") %>%
  kable_styling()
```

```{r animation-add-cities}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            hjust = -0.2,
            vjust = 1) +
  geom_point(data = sakura_overall,
             aes(longitude, latitude)) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15)
```

At this point, the animation is basically done! From here, we are tweaking to make it clearer for the our audience and make it look nice.



First thing we'll do is replace the points representing flower and bloom dates with a sakura icon. I bought mine from iconfinder.com, but you can find free, open-license icons out there.

To do this, we'll use the `{ggimage}` package, which adds `geom_image()`. This acts just like `geom_point()`, except we can point it to an image file and it will display that instead of a point.

I had to do a bunch of fiddling to get the size the way I wanted it.

```{r animation-add-sakura-image}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "Day of year: {round(frame_time, 0)}")

animate(p, start_pause = 15, end_pause = 15)
```


Convert day of year to calendar date.

This step is a bit of a hack. If there are better ways to do it, I'd love to learn them. But this method, though ugly, at least works.

We want to convert day of year into a simple calendar date, like "January 1" or "April 15". (Apologies to my non-North American friends for how I'm formatting the date!)

- Converting day of year into date by adding the `frame_time`, which uses day of the year, to a "base date". Here, we use "2019-01-01". We need a year because date objects need to be a specific date, which requires year. And we use 2019 -- instead of 2020, when this post was written -- because 2020 is a leap year and we don't to include that extra day in February.
- Converting that date into a format that shows only month and day (technically, the labels are in 2019, but we're just hiding the year so that we get a more generic calendar date)

```{r animation-convert-to-date}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "{ymd(\"2019-01-01\") + frame_time}")

animate(p, start_pause = 15, end_pause = 15)
```

```{r animation-convert-to-date-no-year}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}")

animate(p, start_pause = 15, end_pause = 15)
```


Change font. If you haven't used `{extrafont}` before, you'll need to do some setup (GitHub README: https://github.com/wch/extrafont).

I used the Manga Temple font (https://www.dafont.com/manga-temple.font) by Blambot (https://twitter.com/blambot). Note that it's free for personal use, but not for commercial use. Full licence details are here: https://blambot.com/pages/licenses

```{r animation-change-font}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "grey95",
          colour = "grey50") +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1) +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple"))

animate(p, start_pause = 15, end_pause = 15)
```


Change colour of land, background, and font. Going for a creamy/brown vibe.

```{r animation-change-colour}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy,
                    end = full_bloom_doy,
                    enter_length = 5,
                    exit_length = 5) +
  labs(title = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA))

animate(p, start_pause = 15, end_pause = 15)
```



Add growing/shrinking sakura, using `enter_grow()` and `exit_shrink()`. Also using `exit_drift()` with its `y_mod` argument so that they "fall" and shrink simultaneously as they exit. `exit_*()` functions can be combined to get neat behaviours like this.

```{r add-growing-shrinking-transition}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA))

animate(p, start_pause = 15, end_pause = 15)
```



Slow overall animation speed.

```{r animation-slow-speed}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(text = element_text(family = "Manga Temple", colour = "#3D1308"),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA))

animate(p, nframes = 151 * 2, start_pause = 15, end_pause = 15)
```



Add title.

Title text and basic formatting of size, etc.

```{r animation-add-title}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.03) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "When do cherry blossoms bloom in Japan?",
       subtitle = paste("Hanami (\"flower viewing\") is a Japanese tradition over 1,000 years old, usually associated with",
                     "cherry blossom trees. The sakura zensen (cherry blossom front) shows the typical date of",
                     "flowering and full bloom across Japan.",
                     "\n"),
       sep = "\n",
       caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA),
        plot.title = element_textbox_simple(family = "Manga Temple", colour = "#3D1308"),
        plot.subtitle = element_textbox_simple(family = "Manga Temple", colour = "#3D1308"),
        plot.caption = element_text(family = "Manga Temple", colour = "#3D1308"))

animate(p, nframes = 151 * 2, start_pause = 15, end_pause = 15)
```



Adjust title size and location.

Very fiddly step. I often spend a lot of time on this step making sure everything is just right, through trial and error. Specifically mention `halign` and `margin` arguments.

```{r animation-adjust-title-size-location}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.028) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "When do cherry blossoms bloom in Japan?",
       subtitle = paste("Hanami (\"flower viewing\") is a Japanese tradition over 1,000 years old, usually associated with",
                     "cherry blossom trees. The sakura zensen (cherry blossom front) shows the typical date of",
                     "flowering and full bloom across Japan.",
                     "\n"),
       sep = "\n",
       caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA),
        plot.margin = margin(10, 50, 10, 50),
        plot.title = element_textbox_simple(family = "Manga Temple", colour = "#3D1308",
                                            halign = 0.5, margin = margin(10, 10, 10, 10)),
        plot.subtitle = element_textbox_simple(family = "Manga Temple", colour = "#3D1308", halign = 0.5),
        plot.caption = element_text(family = "Manga Temple", colour = "#3D1308"))

animate(p, nframes = 151 * 2, start_pause = 15, end_pause = 15)
```



Add colour to title using `ggtext`.

The `<span>` tags are intimidating, but you get used to them. `ggtext` overview here: https://github.com/wilkelab/ggtext

```{r animation-add-colour-to-title}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 1,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.028) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "When do <span style = 'color:#D47FA8'>**cherry blossoms**</span> bloom in Japan?",
       subtitle = paste("<span style = 'color:#D47FA8'>**Hanami**</span> (\"flower viewing\") is a Japanese tradition over 1,000 years old, usually associated with",
                     "cherry blossom trees. The <span style = 'color:#D47FA8'>**sakura zensen**</span> (cherry blossom front) shows the typical date of",
                     "flowering and full bloom across Japan.",
                     "\n"),
       sep = "\n",
       caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA),
        plot.margin = margin(10, 50, 10, 50),
        plot.title = element_textbox_simple(size = 11, family = "Manga Temple", colour = "#3D1308",
                                            halign = 0.5, margin = margin(10, 10, 10, 10)),
        plot.subtitle = element_textbox_simple(size = 8, family = "Manga Temple", colour = "#3D1308", halign = 0.5),
        plot.caption = element_text(size = 14, family = "Manga Temple", colour = "#3D1308"))

animate(p, nframes = 151 * 2, start_pause = 15, end_pause = 15)
```



```{r final-for-now-animation-brown}
p <- ggplot() +
  borders(regions = "Japan",
          fill = "#AA9C8F",
          colour = NA) +
  geom_point(data = sakura_labels,
             aes(longitude, latitude),
             size = 2,
             colour = "#3D1308") +
  geom_text(data = sakura_labels,
            aes(longitude, latitude, label = station_name),
            size = 8,
            colour = "#3D1308",
            family = "Manga Temple",
            hjust = -0.2,
            vjust = 1) +
  geom_image(data = sakura_overall,
             aes(longitude, latitude, image = "sakura-no-border.svg"),
             size = 0.028) +
  coord_map() +
  theme_map() +
  transition_events(start = flower_doy - 3,
                    end = full_bloom_doy,
                    range(0, 151),
                    enter_length = 6,
                    exit_length = 6) +
  enter_grow() +
  exit_drift(y_mod = -1) +
  exit_shrink() +
  labs(title = "When do <span style = 'color:#D47FA8'>**cherry blossoms**</span> bloom in Japan?",
  subtitle = paste("<span style = 'color:#D47FA8'>**Hanami**</span> (\"flower viewing\") is a Japanese tradition over 1,000 years old, usually associated with",
                     "cherry blossom trees. The <span style = 'color:#D47FA8'>**sakura zensen**</span> (cherry blossom front) shows the typical date of",
                     "flowering and full bloom across Japan.",
                     "\n"),
                     sep = "\n",
  caption = "{paste(month(ymd(\"2019-01-01\") + frame_time, label = TRUE, abbr = FALSE),
                  day(ymd(\"2019-01-01\") + frame_time))}") +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "#FFF9F3", colour = NA),
        plot.margin = margin(10, 50, 10, 50),
        plot.title = element_textbox_simple(size = 32, family = "Manga Temple", colour = "#3D1308",
                                            halign = 0.5, margin = margin(10, 10, -40, 10)),
        plot.subtitle = element_textbox_simple(size = 20, family = "Manga Temple", colour = "#3D1308", halign = 0.5),
        plot.caption = element_text(size = 32, family = "Manga Temple", colour = "#3D1308"))

anim <- animate(p, nframes = 151 * 2, start_pause = 15, end_pause = 15, height = 1600, width = 1100)

### Save animation
### anim_save("sakura-front-brown.gif", animation = anim)
```
